---
title: "Linear regression Bike sharing"
author: "Krishna P Koirala"
date: "5/29/2018"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
bike <- read.csv("bikeshare.csv")
```

## Including Plots

You can also embed plots, for example:

```{r}
head(bike)
```

# Targed variable is count

# EDA

```{r}
library(ggplot2)
ggplot(aes(temp, count), data = bike) + geom_point(alpha = .3, aes(color = temp)) + theme_bw()
```

# Convert to time stamp


```{r}
bike$datetime <- as.POSIXct(bike$datetime)
class(bike$datetime)
```


```{r}
p1 <- ggplot(aes(datetime, count), data = bike) + geom_point(aes(color = temp), alpha = 0.5)
p1
```


```{r}
p1 + scale_color_continuous(low = '#55D8CE', high = '#FF6E2E') + theme_bw()
```

```{r}
cor(bike$temp, bike$count)
```

```{r}
ggplot(aes(factor(season), count), data = bike) + geom_boxplot(aes(color = factor(season))) + theme_dark()
```


# Feature Engineering, creating hour column


```{r}
bike$hour <- sapply(bike$datetime, function(x){format(x, "%H")})
head(bike)
```

# Scatterplot

```{r}
library(dplyr)
# filter()= subset(), filter() only works with dplyr library 
p2 <- ggplot(aes(hour, count), data = filter(bike, workingday == 1)) + geom_point(aes(color = temp), alpha = .3, position = position_jitter(w= .4, h = 0)) + theme_bw()
p2
```


```{r}
p2 <- p2 + scale_color_gradientn(colours = c('dark blue', 'blue', 'light blue', 'light green', 'yellow', 'orange', 'red'))
p2
```


```{r}
p3 <- ggplot(aes(hour, count), data = filter(bike, workingday == 0)) + geom_point(aes(color = temp), alpha = .2, position = position_jitter(w= .4, h = 0)) + theme_bw()
p3
```


```{r}
p3 <- p3 + scale_color_gradientn(colours = c('dark blue', 'blue', 'light blue', 'light green', 'yellow', 'orange', 'red'))
p3
```


# Model Building

This is time series data thats why cannot use all variables to predict
count column. Just using temp column

```{r}
temp.model <- lm(count ~ temp, bike)
```

```{r}
summary(temp.model)
```

# How many bike rental counts at 25 c?

```{r}
# One way of answering above question
mx_plus_c <- 9.1705*25 + 6.0462
mx_plus_c
```

```{r}
# The othe way
temp.test <- data.frame(temp = c(25))
predict(temp.model, temp.test)
```

# changing hour column to numeric

```{r}
bike$hour <- sapply(bike$hour, as.numeric)
```

# Another model

```{r}
# creating model except some columns
model <- lm(count ~ . -casual - registered - datetime - atemp, bike )
summary(model)
```

This model woul not be so good to predict time series analysis.

